Coding theory package for {\it Macaulay2}

### 2021-05-23 tex/paper.tex

\begin{abstract}
In this {\it Macaulay2} \cite{M2} package we implement an object called {\tt LinearCode}. We implement functions that compute basic parameters and objects associated with a linear code, such as generator and parity check matrices, the dual code, length, dimension, and minimum distance, among others. We implement an object called {\tt EvaluationCode}, a construction which allows to study linear codes using tools of
algebraic geometry and commutative algebra. We implement functions to generate important families of linear codes such as Hamming codes, cyclic codes, Reed--Solomon codes, Reed--Muller codes, Cartesian codes, monomial--Cartesian codes, and toric codes. In addition, we implement functions for the syndrome decoding algorithm and locally recoverable code construction, which are important tools in applications of linear codes. The package \textit{CodingTheory} is available at \url{https://github.com/Macaulay2/Workshop-2020-Cleveland/tree/CodingTheory/CodingTheory}
\end{abstract}

@@

\begin{minipage}[t]{0.53\textwidth}
\begin{itemize}[leftmargin=0.55 cm]
  \item {\tt n, k, d, m, s}, positive integers
  \item {\tt F = GF(q)}, a finite field with {\tt q} elements
  \item {\tt g(x)}, a polynomial in {\tt F[x]}
 \end{itemize}
\end{minipage}
\begin{minipage}[t]{0.45\textwidth}
 \begin{itemize}[leftmargin=0.55 cm]
  \item {\tt E}, a list of elements of {\tt F}
  \item {\tt L}, a list of vectors in $F^{\mathtt{m}}$
 \end{itemize}
\end{minipage}

@@

\bibitem{Vardy} A. Vardy, "The intractability of computing the minimum distance of a code," in IEEE Transactions on Information Theory, vol. 43, no. 6, pp. 1757-1766, Nov. 1997, doi: 10.1109/18.641542.

\bibitem{carvalho4} C. Carvalho, V. G. Neumann and H. H. L\'opez,
Projective Nested Cartesian Codes,
Bull Braz Math Soc, New Series (2016).

\bibitem{CSTVV} S. M. Cooper, A. Seceleanu, S. O. Toh\v{a}neanu,
M. Vaz Pinto and R. H. Villarreal, 
Generalized minimum distance functions and algebraic invariants of
Geramita ideals, Adv. in Appl. Math. {\bf 112} (2020), 101940.

@@

\bibitem{GLS} L. Gold, J. Little, H. Schenck,
{Cayley-Bacharach and evaluation codes on complete intersections},
 J. Pure Appl. Algebra  {\bf 196} (1) (2005) 91--99.

@@

\bibitem{Ha1} J. Hansen, { Toric Surfaces and Error--correcting
    Codes} in Coding Theory, Cryptography, and Related Areas, Springer
 (2000), pp. 132-142.

@@

\bibitem{LSc} J. Little, H. Schenck,
{\em Toric surface codes and Minkowski sums},
SIAM J. Discrete Math. {\bf  20}  (2006),  no. 4, 999--1014 (electronic).

\bibitem{lopez-villa} H. H. L\'opez, C. Renter\'ia-M\'arquez and R. H. Villarreal,
Affine Cartesian codes,
Des.\ Codes Cryptogr. \textbf{71}(1) (2014) 5--19.

@@

\bibitem{MPV} J. Mart\'inez-Bernal, Y. Pitones and R. H. Villarreal,
Minimum distance functions of graded ideals and Reed-Muller-type codes,
J. Pure Appl. Algebra 221 (2017), 251--275.

\bibitem{MPV2} J. Mart\'inez-Bernal, Y. Pitones and R. H. Villarreal,
Minimum distance functions of complete intersections,
J. Algebra Appl. 17 (2018), no. 11, 1850204 (22 pages).

\bibitem{algcodes} C. Renter\'\i a, A. Simis and R. Villarreal,
Algebraic methods for parameterized codes and invariants of vanishing ideals over finite fields,
Finite Fields Appl. {\bf 17} (2011), no. 1, 81--104.

\bibitem{Ru}D. Ruano, 
{On the parameters of $r$-dimensional toric codes},
Finite Fields and Their Applications {\bf 13} (2007), pp. 962--976.

@@

\bibitem{SoSo} I.~Soprunov, J.~Soprunova, {Toric surface codes and Minkowski length of polygons}, SIAM J. Discrete Math. {\bf 23}, Issue 1, (2009) 384-400

\bibitem{So1} I. Soprunov,  {Toric complete intersection codes.} Journal of Symbolic Computation. {\bf 50}, (2013) 374--385.

\bibitem{TamoBarg} I. Tamo and A. Barg,
A Family of Optimal Locally Recoverable Codes,
IEEE Transactions on Information Theory {\bf 60} (2014), no. 8, 4661--4676.

\bibitem{renteria-tapia-ca2} C. Renter\'\i a and H. Tapia-Recillas, 
Reed-Muller codes: an ideal theory approach, Comm. Algebra {\bf 25}
(1997), no. 2, 401--413.

\bibitem{van-lint} J. H. Van Lint,
{ Introduction to coding theory},
Third edition, Graduate Texts in Mathematics {\bf 86}, Springer-Verlag, Berlin, 1999.

\bibitem{monalg} R. H. Villarreal, {\it Monomial Algebras\/},
Second edition, 
Monographs and Research Notes in Mathematics, Chapman and Hall/CRC,
Boca Raton, FL, 2015.
\end{thebibliography}
\end{document}

>>>>

\begin{abstract}
In this {\it Macaulay2} \cite{M2} package we implement an object called {\tt LinearCode}. We implement functions that compute basic parameters and objects associated with a linear code, such as generator and parity check matrices, the dual code, length, dimension, and minimum distance, among others. We implement an object called {\tt EvaluationCode}, a construction which allows to study linear codes using tools of
algebraic geometry and commutative algebra. We implement functions to generate important families of linear codes such as Hamming codes, cyclic codes, Reed--Solomon codes, Reed--Muller codes, Cartesian codes, monomial--Cartesian codes, and toric codes. In addition, we implement functions for the syndrome decoding algorithm and locally recoverable code construction, which are important tools in applications of linear codes.% The package \textit{CodingTheory} is available at \url{https://github.com/Macaulay2/Workshop-2020-Cleveland/tree/CodingTheory/CodingTheory}
\end{abstract}

@@

\begin{minipage}[t]{0.53\textwidth}
\begin{itemize}[leftmargin=0.55 cm]
  \item {\tt n, k, d, m, s}, positive integers
  \item {\tt F = GF(q)}, a finite field with {\tt q} elements
  \item {\tt g(x)}, a polynomial in {\tt F[x]}
 \end{itemize}
\end{minipage}
\begin{minipage}[t]{0.45\textwidth}
 \begin{itemize}[leftmargin=0.55 cm]
  \item {\tt E}, a list of elements of {\tt F}
  \item {\tt L}, a list of vectors in ${\tt F}^{\mathtt{m}}$
 \end{itemize}
\end{minipage}

@@

\bibitem{carvalho4} C. Carvalho, V. G. Neumann and H. H. L\'opez,
Projective Nested Cartesian Codes,
Bulletin of the Brazilian Mathematical Society, New Series, {\bf 48} (2017), no. 2, 283--302.

\bibitem{CSTVV} S. M. Cooper, A. Seceleanu, S. O. Toh\v{a}neanu, M. Vaz Pinto and R. H. Villarreal, 
Generalized minimum distance functions and algebraic invariants of Geramita ideals,
Advances in Applied Mathematics, {\bf 112} (2020), 101940.

@@

\bibitem{GLS} L. Gold, J. Little, H. Schenck,
Cayley-Bacharach and evaluation codes on complete intersections,
Journal of Pure and Applied Algebra, {\bf 196} (2005), no. 1, 91--99.

@@

\bibitem{Ha1} J. Hansen,
Toric Surfaces and Error--correcting Codes,
in Coding Theory, Cryptography, and Related Areas, (2000), pp. 132-142. Springer, Berlin, Heidelberg.

@@

\bibitem{LSc} J. Little and H. Schenck,
Toric surface codes and Minkowski sums,
SIAM Journal on Discrete Mathematics, {\bf  20}  (2006),  no. 4, 999--1014.

\bibitem{lopez-villa} H. H. L\'opez, C. Renter\'ia-M\'arquez and R. H. Villarreal,
Affine Cartesian codes,
{Designs, Codes and Cryptography}, {\bf 71} (2014), no. 1, 5--19.

@@

\bibitem{MPV2} J. Mart\'inez-Bernal, Y. Pitones and R. H. Villarreal,
Minimum distance functions of complete intersections,
Journal of Algebra and Its Applications, {\bf 17} (2018), no. 11, 1850204 (22 pages).

\bibitem{MPV} J. Mart\'inez-Bernal, Y. Pitones and R. H. Villarreal,
Minimum distance functions of graded ideals and Reed-Muller-type codes,
Journal of Pure and Applied Algebra. {\bf 221} (2017), no. 2, 251--275.

\bibitem{algcodes} C. Renter\'\i a, A. Simis and R. Villarreal,
Algebraic methods for parameterized codes and invariants of vanishing ideals over finite fields,
Finite Fields and Their Applications, {\bf 17} (2011), no. 1, 81--104.

\bibitem{Ru}D. Ruano, 
On the parameters of $r$-dimensional toric codes,
Finite Fields and Their Applications, {\bf 13} (2007), no. 4, pp. 962--976.

@@

\bibitem{SoSo} I.~Soprunov, J.~Soprunova,
Toric surface codes and Minkowski length of polygons,
SIAM Journal on Discrete Mathematics, {\bf 23} (2009), no. 1, 384--400.

\bibitem{So1} I. Soprunov,
Toric complete intersection codes,
Journal of Symbolic Computation, {\bf 50} (2013), 374--385.

\bibitem{TamoBarg} I. Tamo and A. Barg,
A Family of Optimal Locally Recoverable Codes,
IEEE Transactions on Information Theory, {\bf 60} (2014), no. 8, 4661--4676.

\bibitem{renteria-tapia-ca2} C. Renter\'\i a and H. Tapia-Recillas, 
Reed-Muller codes: an ideal theory approach,
Communications in Algebra, {\bf 25} (1997), no. 2, 401--413.

\bibitem{van-lint} J. H. Van Lint,
Introduction to coding theory,
Third edition, Graduate Texts in Mathematics {\bf 86}, Springer-Verlag, Berlin, 1999.

\bibitem{Vardy} A. Vardy,
The intractability of computing the minimum distance of a code,
IEEE Transactions on Information Theory, {\bf 43} (1997), no. 6, 1757--1766.

\bibitem{monalg} R. H. Villarreal, {\it Monomial Algebras\/},
Second edition, 
Monographs and Research Notes in Mathematics, Chapman and Hall/CRC,
Boca Raton, FL (2015).
\end{thebibliography}
\end{document}
